---
title: 'Name: Loubna Sasso - nyc SQF'
author: "| Student number: 20167595 \n"
date: "`r format(Sys.time(), '%X, %d %B, %Y')`"
output:
  word_document: default
  pdf_document: default
  html_document: default
editor_options:
  markdown:
    wrap: 72
---

# Originality declaration

I, \[**Loubna Sasso**\], confirm that the work presented in this
assessment is my own. Where information has been derived from other
sources, I confirm that this has been indicated in the work.

date: `r format(Sys.time(), '%d %B, %Y')`

# Start your response here

## 01 Initial project scope

SQF policy has been the subject of racial profiling over the years,
therefore analyzing where SQF polcies are being implemented across
different areas of NYC and who they are targeted towards is essential in
order to help overcome racial profiling by informing policies for
specific precincts to ensure equality. Moreover Research shows that
"persons of African and Hispanic descent were stopped more frequently
than white people, even after controlling for precinct variability and
race-specific estimates of crime participation." Therefore analyzing
where these differences occur in different parts on NYC can be
insightful to implement necessary changes.

The databasets used for this analysis comes from NYC open data and New
York City Police Department. The New York City Police Department Dataset
contains 9545 SQF entries contains entries for and 83 different fields.
The dataset includes important information on the various SQF throughout
2020, at different months, and time of day as well as the race and
gender of the people being questioned.

The fields that will be used for this analysis include the latitude and
longitude coordinates, suspect race description and, month. I will
analyse the pattern of sqf on black people over the summer period
(june,july,august,spetember). This process will repeated for 'white
people' if i have time during the 6 hours given, and a comparison
between sqf of black vs white people will be made

#### Research Question:

What patterns off risk (SQF) can be observed in the different precincts
in NYC for black people and how might this pattern change in other
racial groups?

#### Hypothesis:

HYPOTHESIS 1 h0: There are no noticeable spacial cluster pattern of SQF
policy that can be observed for the black people across NYC h1: There
are noticeable spacial cluster pattern of SQF policy that can be
observed for the black people across NYC

HYPOTHESIS 2 (moran, getis gerrys) Ho: There are no similarities in
spatial autocorrelation of (SQF) policy between black and white people\
H1: There are similarities in spatial autocorrelation of (SQF) policy
between black and white people

#### Data Limitations

-   Point data also lie outside the boundaries of nyc boundaries that
    could influence what's happening inside
-   Is it ethical to include peoples names in the data provided?

#### Analysis Assumptions and focus areas

-   Assume that points outside NYC boundary have no influence on the
    outcome of data inside NYC
-   Focus on black community and then possibly white community if i have
    time
-   Assume data is complete regardless of na values

#### Logical Plan

Preprocess the data: - load in data, filter data to black people being
questioned and transform it - transform lat and long values to point
data, make all points distinct, clip all point to NYC boundary - st_join
points to spatial data: group by, summarize and get a density map

Analyzing the data: - Descriptive statistics of all SQF on black people
density data: BOX PLOT - Explore clustering of SQF using Point pattern
analysis: kerney, ripleys and DBSCAN - Explore different Global methods
to analyze levels Spatial autocorrelation of SQF in NYC - Explore
different local methods to analyze clustering across communities in NYC

## 02 Load in libraries

```{r}
library(spatstat)
library(here)
library(sp)
library(rgeos)
library(maptools)
library(GISTools)
library(tmap)
library(sf)
library(geojson)
library(geojsonio)
library(tmaptools)
library(stringr)
library(tidyverse)
library(raster)
library(fpc)
library(dbscan)
library(ggplot2)
library(OpenStreetMap)
library(janitor)
library(dplyr)
library(spdep)

library(plotly)
library(rgdal)
library(broom)
library(mapview)
library(crosstalk)
library(car)
library(fs)
library(tidypredict)
library(corrr)
library(spatialreg)
library(spdep)
library(spgwr)
library(joineR)
library(units)
library(ggplot2)
library(xlsx)
library(readxl)


```

## 03 Load in data +transform Coordinates

```{r}

#read in Police data and transform it
police <- st_read(here::here ("Data", "Police Precincts.geojson")) %>%
  st_transform(., 32618) #2908


#read in sqf data, remove all na values and clean names
sqf  <- read_excel(here::here ("Data", "sqf-2020.xlsx"))%>%
  clean_names()

```

## 04 Initianl Data Cleaning and wrangling

for sql: -select fields: suspect_race_description, stop_location_x,
stop_location_x, suspect_arrest_offense, month2 -filter: Black, white,
white hispanic -drop n/a values -transform

```{r}

#1) for sqf csv data i want to:

#select columns you want. 
sqf <-  sqf%>%
  dplyr::select(suspect_race_description, stop_location_x, stop_location_y, month2)


# first filter only black and drop na vals in lat and long
sqf_black <-  sqf%>%
  filter(suspect_race_description == "BLACK")%>%
  filter(month2 == "June" | month2 == "July" | month2 == "August" | month2 == "September" )%>%
  drop_na(stop_location_x)%>%
  drop_na(stop_location_y)
                  
  
# first filter only white and drop any na vals in lat and long. this will be used later to compare if i have time
sqf_white <-  sqf%>%
  filter(suspect_race_description == "WHITE HISPANIC" | suspect_race_description == "WHITE")%>%
  filter(month2 == "June" | month2 == "July" | month2 == "August" | month2 == "September" )%>%
  drop_na(stop_location_x)%>%
  drop_na(stop_location_y)

  

#2) transform csv sqf_black to point data: st_as sf. 
sqf_black_sf <- sqf_black%>%
  st_as_sf(.,coords = c ( "stop_location_x", "stop_location_y"), crs=2908)%>%
  st_transform(.,32618)

  
#make all points distinct 
sqf_black_sf <- sqf_black_sf%>%
  distinct(geometry, .keep_all=T)


#3) overlay point data to police districts and clip if required
tmap_mode("view")
tm_shape(police) +
  tm_polygons(col = NA, alpha = 0.5) +
tm_shape(sqf_black_sf) +
  tm_dots(col = "blue")


#clipping point to xxx map boundary
sqf_black_sf <- sqf_black_sf [police,]

#plot again to see if they are clipped
tmap_mode("view")
tm_shape(police) +
  tm_polygons(col = NA, alpha = 0.5) +
tm_shape(sqf_black_sf) +
  tm_dots(col = "blue")

```

## 05 Density map and Descriptive Statistics: sqf black

lets get an idea of our sqf black data density and distribution

```{r}

#1) observe areas of highest and lowest density of sqf-black
#could similar  densities in adjacent districts mean there is some degree of clustering happening in these areas?


#st_join point data to the police map to allow for continuous data observation
police_sqf_black <- police%>%
  st_join(sqf_black_sf)

#count no. of points in each precinct and add area and density 
police_sqf_black <- police_sqf_black%>%
  add_count(precinct)%>% #counts the no. of points in each precinct (n)
  janitor::clean_names()%>%
  mutate(area = st_area(.))%>% #get area
  mutate (density = n/area)%>%  #get density
  dplyr::select(density, area, precinct, n) #narror down the colomns you want


#group by borough
police_sqf_black <-  police_sqf_black%>%
  group_by(precinct)%>%
  summarise(density = first(density), precinct = first(precinct), area = first(area), 
            count = first(n))



#quick choropleth map based on DENSITY  
tm_shape(police_sqf_black) +
    tm_polygons("density",
        style="jenks",
        palette="PuOr",
        midpoint=NA,
        popup.vars=c("precinct", "density"),
        title="SQF-Black people Density")



#2) Box plot

#first make sure density col is numeric
police_sqf_black$density_numeric <- as.numeric(police_sqf_black$density)

#boxplot(police_sqf_black$density_numeric)
boxplot(police_sqf_black$density_numeric,
main = "SQF-Black people Density Distribution ",
xlab = "density ",
ylab = "",
col = "orange",
border = "brown",
horizontal = TRUE
)


```

Highest levels of sqf amongst black people are indicated in dark purple
in precincts across Manhattan, west of Bronx and north of Brooklyn. it
can also be seen that precincts closer together follow more or less
similar sqf of black people density which could indicate similarity in
governance in these precincts that are influenced by neighbouring
precincts, therefore spatial autocorrelation and clustering occurance.

We can also see the density distribution is positively skewed, meaning
that most of the precints in Nyc exhibit lower densities in of sqf
amongst black people, however there are still a few areas with high
densities. Lets explore further the spatial patterns of these density
distribution across the different precinct through point pattern
analysis and later morans I

##06 Cluster Analysis: Point pattern Analysis:

ANalysis process: 1. create sp and sf objects 2. Density kernel (which
confirms the density map created above) 3. Ripley K (over quadrat)
analysis to test clustering present 4. DBSCAN to show where clusters are
happening

1.Create sp and ppp objects from the sf point data in order to proceed
with ripley k and DBSCAN

```{r}
#1) now set a window as the precinct nyc boundary 
window <- as.owin(police)
plot(window)


#2A)  Create a sp object for eviction (sf to sp)
sqf_black_sp <- sqf_black_sf %>%
  as(., 'Spatial')

#2B) Create a ppp object
sqf_black_ppp <- ppp(x=sqf_black_sp@coords[,1],
                          y=sqf_black_sp@coords[,2],
                          window=window)

#sqf_black_sp@coords[,1]



#3) have a look at the ppp object

sqf_black_ppp %>%
  plot(.,pch=16,cex=0.5, 
       main="SQF Black people")

```

### kernel Density:

This confirms the density map produced earlier.

```{r}
#uses ppp data 
sqf_black_ppp %>%
  density(., sigma=500) %>%
  plot()

```

We can see from this map that the that the precincts that fall in
manhattan , parts of Bronx shows highest densities of SQF concentration,
similar to the first density diagram produced. Now lets explore further
if the points exhibit clustering using ripleys k.

### Ripleys k

"To determining whether a pattern exists in the data is to compare the
observed distribution of points to an idealised random distribution. If
the points do not conform to a random distribution, then it is likely
that there is some unobserved factor or factors which could be causing
non-random clustering or dispersal to occur" (Adam Dennett and Sam).

Quadrat analysis does not account for different spatial unit sizes such
the different sized precincts in NYC, therefore to overcome this
limitation, i will jump straight in to a ripleys k test to check if the
SQF cases of black people exhibits clustering, This is done by comparing
the observed distribution of points with the Poisson random model for a
whole range of different distance radii.

```{r}

#using the ppp data, plot the ripleys K
Ripleys_K <- sqf_black_ppp%>%
  Kest(., correction="border") %>%
  plot()


Kval_global <- as.data.frame(Kest(sqf_black_ppp, correction = "border"))
Kval_global


```

From the graph, we can see that for all distances, SQF locations appear
to be clustered up until a distance of approximately 3000. Ripleys k
shows us if we have spatial clusters present in our point data, but it
does not tell us WHERE clusters are occurring. To discover this we can
use DBSAN

### DBSCAN

Results of the Ripley's K analysis earlier shows we are getting
clustering up to a radius of around 3000m, with the largest bulge in the
graph also being at around 3000m radius . Therefore, 3000m is probably a
good place to start and we will begin by searching for clusters of at
least 4. if this distance isnt suitable, i will explore the elbow plot
and use my own judgement to set eps.

Upon attempting the clusters at 3000, it seems to pick up all the
points, therefore i have adjuster the eps distance to 900 as per the
elbow plot. Moreover, having the cluster threshold at 4, you can see
clusters emerge.

```{r}

#first check the coordinate reference system of the this spatial polygon:
st_geometry(police)


#1.extract the points from the 'spatial points (sp)' data frame
sqf_black_sp_extract <- sqf_black_sp%>% #takes 2A above
  coordinates(.)%>%
  as.data.frame()

#2. run the DBSCAN using the extracted sp points: 
#i choose clustering of eviction to be at 900 and min number of 4 points
DB_sqf_black <- sqf_black_sp_extract  %>% 
  fpc::dbscan(.,eps = 900, MinPts = 4) 

#now plot the results
plot(DB_sqf_black, sqf_black_sp_extract , main = "SQF-Black DBSCAN Output", frame = F)
plot(police$geometry, add=T)




#3. lets double check appropriate eps distance with the elbow plot:

sqf_black_sp_extract%>% #takes the point extracted from sp object (1. above)
  dbscan::kNNdistplot(.,k=4)
#shows 4000, howveer this seems too large so i'll keep it at 1000 as selected earlier


```

The outcome of the dbsan confirms that clustering of SQF on black people
seems to follow a similarity in the density distribution mapped earlier.
Using this map, we can not only confirm there there is clusters of SQF
on black people happening, but that the highest clusters are occuring in
the same areas mentioned earlier-manhattan, Brooklyn and Bronx.

Could these precincts be influenced by neighboring areas that could
perhaps explain their clustering? let explore this spatial
autocorrelation using morans I.

## 07 Spatial Autocorrelation: SQF Black people

According to Waldo R. Tobler, "Everything is related to everything else.
But near things are more related than distant things". Lets investigate
precinct-level spatial dependency using localized version of morana's I
in order to see hot spots of where SQF programs on black people are
happening.

If they exibit spatial autocorrelation then this would mean that
precinct spatial units are influenced by other neighboring precincts and
hence may explain potential clusters happening in areas that are
geographically close.If this is the case, then similar policies for
neighbouring precincts that can be adopted.

### Spatial weight matrix

Before performing our clustered analysis, we need to generate a spatial
weights matrix

```{r}
#1. calculate the centroids of all polygons in NYC
centroids_police <- police_sqf_black%>%
  st_centroid()%>%
  st_geometry()

plot(centroids_police)


#2. generate neighbor lists

#queen's case neighbors : doesn't take centroids. taken adjacent neighbors.
NL_queens <- police_sqf_black %>%   
  poly2nb(., queen=T)

#or k nearest neighbors : take centroids. 
NL_KNN <-centroids_police %>%  
  knearneigh(., k=4)

NL_KNN <- NL_KNN %>%
  knn2nb()

#now plot them both: Fist take the NL then the centroid of the spatial object
plot(NL_queens, st_geometry(centroids_police), col="red")
plot(police_sqf_black$geometry, add=T)

plot(NL_KNN, st_geometry(centroids_police), col="blue")
#add a map underneath
plot(police_sqf_black$geometry, add=T)


#3. create a spatial weights matrix object from these weights
# using row standardization 'w'

SWM_queens <- NL_queens %>%
  nb2listw(., style="W") 

SWM_knn <-NL_KNN %>%
  nb2listw(., style="W")
```

#### Local Moran's I

final output map 1:

```{r}

tmap_mode("plot")

#1. Local morans density
local_moran_density <-  police_sqf_black %>%
  pull(density_numeric) %>% #column name from police_sqf_black use the density values or count to get local morans
  as.vector()%>%
  localmoran(., SWM_knn)%>%  #spatial weight matrix of knn (generated from before from spatial data)
  as_tibble()
  

slice_head(local_moran_density, n=5) #view the values


# 2. Copy the I score (column 1) and the z-score sd. (column 4)) back into police_sqf_black data

police_sqf_black <- police_sqf_black %>%
  mutate(density_I =as.numeric(local_moran_density$Ii))%>% #local_moran_density is tibble above, specific col =Ii
  mutate(density_Iz =as.numeric(local_moran_density$Z.Ii))


#3.lets map them out:

#set breaks
breaks1<-c(-1000,-2.58,-1.96,-1.65,1.65,1.96,2.58,1000)

#set colors
MoranColours<- rev(brewer.pal(8, "RdGy"))



tmap_mode("plot")
#plot interactive map
tm_shape(police_sqf_black) +
  tm_layout( title = "Local Moran's I- 
SQF:Black People")+
  tm_layout(frame = FALSE)+
  tm_polygons("density_Iz",
        style="fixed",
        breaks=breaks1,
        palette=MoranColours,
        midpoint=NA,
        title="Local Moran's I",
        legend.hist= TRUE)+
  tm_compass() +
  tm_scale_bar() +
  tm_legend(position=c(1,0.0))+
  tm_credits("(c) New York City Police Department", position=c(0.5,0.0))


```

The limitation of using local morans i alsone is that it is susceptible
to shape and sizes of the spatial units therefore in order to overcome
this issues, this map should be read along side the dbscan map that
shows spatial clusters.

This local morans I map shows rare cases of the highest hot spots of SQF
on black people occurring in precincts: 23,24,26,28 and 30. By reading
this map along with the dbscan cluster analysis, we can see that these
precincts not only exhibit high clusters of SQF cases on black people
but also have high influence on each other.

## 08 White People SQF

-   observe clustering of white people SQF and compare it to black
    people
-   Overlay points of of SQF White people to spatial autocorrelation
    data maps of SQF black
-   Develop local moran's I map for SQF white people data and compare
    the spatial autocorrelation

```{r}

# first filter only white and drop any na vals in lat and long. this will be used later to compare if i have time
sqf_white <-  sqf%>%
  filter(suspect_race_description == "WHITE HISPANIC" | suspect_race_description == "WHITE")%>%
  filter(month2 == "June" | month2 == "July" | month2 == "August" | month2 == "September" )%>%
  drop_na(stop_location_x)%>%
  drop_na(stop_location_y)

  

#2) transform csv sqf_white to point data: st_as sf. 
sqf_white_sf <- sqf_white%>%
  st_as_sf(.,coords = c ( "stop_location_x", "stop_location_y"), crs=2908)%>%
  st_transform(.,32618)

  
#make all points distinct 
sqf_white_sf <- sqf_white_sf%>%
  distinct(geometry, .keep_all=T)


#3) overlay point data to police districts and clip if required
tmap_mode("view")
tm_shape(police) +
  tm_polygons(col = NA, alpha = 0.5) +
tm_shape(sqf_white_sf) +
  tm_dots(col = "blue")


#clipping point to xxx map boundary
sqf_white_sf <- sqf_white_sf[police,]

#plot again to see if they are clipped
tmap_mode("view")
tm_shape(police) +
  tm_polygons(col = NA, alpha = 0.5) +
tm_shape(sqf_white_sf) +
  tm_dots(col = "blue")



```

Density map of white people

```{r}
#1) observe areas of highest and lowest density of sqf-white


#st_join point data to the police map to allow for continuous data observation
police_sqf_white <- police%>%
  st_join(sqf_white_sf)

#count no. of points in each precinct and add area and density 
police_sqf_white <- police_sqf_white%>%
  add_count(precinct)%>% #counts the no. of points in each precinct (n)
  janitor::clean_names()%>%
  mutate(area = st_area(.))%>% #get area
  mutate (density = n/area)%>%  #get density
  dplyr::select(density, area, precinct, n) #narror down the colomns you want


#group by borough
police_sqf_white <-  police_sqf_white%>%
  group_by(precinct)%>%
  summarise(density = first(density), precinct = first(precinct), area = first(area), 
            count = first(n))



#quick choropleth map based on DENSITY  
tm_shape(police_sqf_white) +
    tm_polygons("density",
        style="jenks",
        palette="PuOr",
        midpoint=NA,
        popup.vars=c("precinct", "density"),
        title="SQF-White people Density")

```

we can see some slight differences when compared to density of black
people, however they both show high density in Manhattan. Lets explore
the clusters of SQF white people in order to make further comparisons.

Cluster analysis: White people

ANalysis process: 1. create sp and sf objects 2. Density kernel (which
confirms the density map created above) 3. Ripley K (over quadrat)
analysis to test clustering present 4. DBSCAN to show where clusters are
happening

1.Create sp and ppp objects from the sf point data in order to proceed
with ripley k and DBSCAN

```{r}
#1) now set a window as the precinct nyc boundary 
window <- as.owin(police)
plot(window)


#2A)  Create a sp object for eviction (sf to sp)
sqf_white_sp <- sqf_white_sf  %>%
  as(., 'Spatial')

#2B) Create a ppp object
sqf_white_ppp <- ppp(x=sqf_white_sp@coords[,1],
                          y=sqf_white_sp@coords[,2],
                          window=window)

#sqf_black_sp@coords[,1]



#3) have a look at the ppp object

sqf_white_ppp %>%
  plot(.,pch=16,cex=0.5, 
       main="SQF white people")

```

### kernel Density:

This confirms the density map produced earlier.

```{r}
#uses ppp data 
sqf_white_ppp %>%
  density(., sigma=500) %>%
  plot()

```

We can see from this map that the similar to SQF black people cases, the
precincts that fall in manhattan and part of Bronx also shows highest
densities of SQF concentration for white people

### Ripleys k

```{r}

#using the ppp data, plot the ripleys K
Ripleys_K <- sqf_white_ppp%>%
  Kest(., correction="border") %>%
  plot()


Kval_global <- as.data.frame(Kest(sqf_black_ppp, correction = "border"))
Kval_global


```

From the graph, we can see that up to a distance of 2100, SQF locations
for white people appear to be clustered and from about 2100 to 2400 they
are dispersed.. Lets perform dbsan

### DBSCAN

i'll begin with eps of 2100 and adjust if needed.

Upon attempting the clusters at 2100, it seems to pick up all the
points, therefore i have adjuster the eps distance to 900 (even though
elbow plot shows 500) i order to match the same distance as black
people. This will allows a fair comparison between the two.

```{r}

#first check the coordinate reference system of the this spatial polygon:
st_geometry(police)


#1.extract the points from the 'spatial points (sp)' data frame
sqf_white_sp_extract <- sqf_white_sp%>% 
  coordinates(.)%>%
  as.data.frame()

#2. run the DBSCAN using the extracted sp points: 
#i choose clustering of eviction to be at 100 and min number of 6 points
DB_sqf_white <- sqf_white_sp_extract  %>% 
  fpc::dbscan(.,eps = 900, MinPts = 4) 

#now plot the results
plot(DB_sqf_white, sqf_white_sp_extract , main = "SQF-White DBSCAN Output", frame = F)
plot(police$geometry, add=T)




#3. lets double check appropriate eps distance with the elbow plot:

sqf_white_sp_extract%>% #takes the point extracted from sp object (1. above)
  dbscan::kNNdistplot(.,k=4)
#shows 4000, howveer this seems too large so i'll keep it at 1000 as selected earlier


```

The outcome of the dbsan confirms that clustering of SQF on white people
seems

fOR the same eps distances, SQF clusters of black people are more
apparent in the precincts that fall inside Brooklyn, However there is a
similarity in clustering of both black and white SQF cases in Manhattan.

Lets explore the spatial autocorrelation of sqf white people to make a
comparison.

### Local Morans

Spatial weight matrix: Before performing our clustered analysis, we need
to generate a spatial weights matrix

```{r}
#1. calculate the centroids of all polygons in NYC
centroids_police_w <- police_sqf_white%>%
  st_centroid()%>%
  st_geometry()

plot(centroids_police_w)


#2. generate neighbor lists

#queen's case neighbors : doesn't take centroids. taken adjacent neighbors.
NL_queens_W <- police_sqf_white %>%   
  poly2nb(., queen=T)

#or k nearest neighbors : take centroids. 
NL_KNN_w <-centroids_police_w %>%  
  knearneigh(., k=4)

NL_KNN_w <- NL_KNN_w %>%
  knn2nb()

#now plot them both: Fist take the NL then the centroid of the spatial object
plot(NL_queens_W, st_geometry(centroids_police_w), col="red")
plot(police_sqf_white$geometry, add=T)

plot(NL_KNN_w, st_geometry(centroids_police_w), col="blue")
#add a map underneath
plot(police_sqf_white$geometry, add=T)


#3. create a spatial weights matrix object from these weights
# using row standardization 'w'

SWM_queens_w <- NL_queens_W %>%
  nb2listw(., style="W") 


SWM_knn <-NL_KNN %>%
  nb2listw(., style="W")

```

#### Local Moran's I

final output map 2:

```{r}

tmap_mode("plot")

#first make sure density col is numeric
police_sqf_white$density_numeric <- as.numeric(police_sqf_white$density)



#1. Local morans density
local_moran_density_w <-  police_sqf_white %>%
  pull(density_numeric) %>% #column name from police_sqf_black use the density values or count to get local morans
  as.vector()%>%
  localmoran(., SWM_knn)%>%  #spatial weight matrix of knn (generated from before from spatial data)
  as_tibble()
  

slice_head(local_moran_density_w, n=5) #view the values


# 2. Copy the I score (column 1) and the z-score sd. (column 4)) back into police_sqf_black data

police_sqf_white <- police_sqf_white %>%
  mutate(density_I =as.numeric(local_moran_density_w$Ii))%>% #local_moran_density is tibble above, specific col =Ii
  mutate(density_Iz =as.numeric(local_moran_density_w$Z.Ii))


#3.lets map them out:

#set breaks
breaks1<-c(-1000,-2.58,-1.96,-1.65,1.65,1.96,2.58,1000)

#set colors
MoranColours<- rev(brewer.pal(8, "RdGy"))



tmap_mode("plot")
#plot interactive map
tm_shape(police_sqf_white) +
  tm_layout( title = "Local Moran's I- 
SQF:white People")+
  tm_layout(frame = FALSE)+
  tm_polygons("density_Iz",
        style="fixed",
        breaks=breaks1,
        palette=MoranColours,
        midpoint=NA,
        title="Local Moran's I",
        legend.hist= TRUE)+
  tm_compass() +
  tm_scale_bar() +
  tm_legend(position=c(1,0.0))+
  tm_credits("(c) New York City Police Department", position=c(0.5,0.0))


```

final output map 3:

```{r}


#set breaks
breaks1<-c(-1000,-2.58,-1.96,-1.65,1.65,1.96,2.58,1000)

#set colors
MoranColours<- rev(brewer.pal(8, "RdGy"))


#map SQF: black people
tmap_mode("plot")
#plot interactive map
tm1 <-  tm_shape(police_sqf_black) +
  tm_layout( title = "Local Moran's: 
Black People")+
  tm_polygons("density_Iz",
        style="fixed",
        breaks=breaks1,
        palette=MoranColours,
        midpoint=NA,
        title="Local Moran's I SQF:Black People")+
  tm_legend(show=FALSE)+
  tm_layout(frame=FALSE)



#map SQF: white people
tmap_mode("plot")
#plot interactive map
tm2 <- tm_shape(police_sqf_white) +
  tm_layout( title = "Local Moran's:
White People")+
  tm_polygons("density_Iz",
        style="fixed",
        breaks=breaks1,
        palette=MoranColours,
        midpoint=NA,
        title="Local Moran's I")+
  tm_legend(show=FALSE)+
  tm_layout(frame=FALSE)



legend1 <- tm_shape(police_sqf_black) +
    tm_polygons("density_Iz",
        style="fixed",
        breaks=breaks1,
        palette=MoranColours,
        midpoint=NA,
        title="SQF:Black people",
        legend.hist= FALSE) +
    tm_scale_bar(position=c(0.2,0.04), text.size=0.6)+
    tm_compass(north=0, position=c(0,0.1))+
    tm_layout(legend.only = TRUE, legend.position=c(0.5,0.25),asp=0.1)+
    tm_credits("(c) New York City Police Department", position=c(0.0,0.0))



#map the legend
legend <- tm_shape(police_sqf_white) +
    tm_polygons("density_Iz",
        style="fixed",
        breaks=breaks1,
        palette=MoranColours,
        midpoint=NA,
        title="SQF:White people",
        legend.hist= FALSE) +
    tm_layout(legend.only = TRUE, legend.position=c(0.5,0.25),asp=0.1)
    




#plot them side by side
t=tmap_arrange(tm1, tm2, legend1, legend, ncol=2)


t



```

Comparing the 2 maps we can see that some precincts exhibit similarities
in levels of autocorrelation specifically in staten island (1.6-2) and
manhattan. In general we can see high levels of spatial autocorrelation
in both maps in precincts inside manhattan which could indicate that
although theyr are both high, there isnt necessarily racial-profiling
happening in manhattan.

## 09 Reflection

The analysis conducted lead us to reach a conclusion that answers the
research question and both hypothesis

Ripleys k indicated spatial clustering was present In both cases of
black people and white people.

DBSAN was used to observe clustered pattern and local morans to
determine hot spots of spatial autocorrelation. The analysis was first
done for all SQF on suspect black people and results showed that
clustering was most present in precincts across Manhattan, west of Bronx
and north of Brooklyn. The clustering performed on SQF of white cases
showed that for a similar eps value, there were similarities in
clustering in Manhattan, however sqf clustering of black people were
higher in Brooklyn which indicates there may be possibilities of racial
profiling of black people in Brooklyn

This Lead us to confirm H1 for the first hypothesis:There are noticeable
spacial cluster pattern of SQF policy that can be observed for the black
people across NYC

Moreover, by cross comparing Morans i of both black and white sqf cases
showed some precincts exhibit similarities in spatial autocorrelation,
with highest levels of spatial autocorrelation for both maps in
precincts inside manhattan. This could indicate there there may be other
factors such internal governance within these precincts that are causing
such strong spatial similarities between these precincts.

This confirms H1 in hypothesis 2: H1: There are are similarities in
spatial autocorrelation of (SQF) policy between black and white people

Limitations to the analysis - Ripley's k is sensitive to the boundaries
of the study area - Addition spatial autocorrelation was not explored
including gerrys and getis -The eps distance outcome of ripleys K could
not be used as it was too large and so 900m assumption was set for both
morans i cases. -NYPD collected the data so there may have been a
confilict of interest

Policy Making -in areas where there are clusters of SQR on black people
such as the bronx, policies to ensure that there is no police violence
can be be implemented. This may include internal trainin of police
forces within these precincts to ensure that racial profiling and racial
profiling does not occur.

Further Exploration -clusters over different periods from 2018 compared
to 2020 to compare changes after black lives matter. -clusters of socio
economic factors that may influence or be linked to SQF racial cases
-local regression models for each cluster:localized analysis and
therefore custom policy making for precinct that exhibit similarities in
spatial pattern

```{r}



library(rmarkdown)
render("exam_response.Rmd", "pdf_document")

```
